Least-Squares Time Reversal of PDEs

### 2022-07-21 adjop/lsrtm/lsrtm.tex
not sure
undiffuse section

Time-reversible wave equations of the form
\begin{align}
  \partial_t u &= \sum_{n \geq 1} c_n \partial^{2n - 1}_x u \label{waveeq}
  \intertext{are trivial cases of this optimization problem, because the quantity $\davg{u^2(x, t)}$ is conserved. 
  This implies that the objective functional's levelsets form sphere's in $u'$-space, because its value is given by $\davg{u'^2(T)} = \davg{u'^2(0)}$.
  The spatial derivatives in (\ref{waveeq}) are anti-self-adjoint (i.e. $L[u] = -L^{\dagger}[u]$), so the path of steepest descent is always in the radial derivation }
  \frac{\delta H}{\delta u(x, 0)} &= -u'(x, 0) \nonumber
  \intertext{In contrast, diffusive equations of the form}
  \partial_t u &= \sum_{n \geq 1} b_n \partial^{2n}_x u \label{waveeq}
  \intertext{do not conserve $\davg{u^2(x, t)}$, and consequently their objective functionals depend on $u'(x, 0)$'s \textit{direction} along with its magnitude. Suppose initial deviation can be written as}
  u'(x, 0) &= \sum_{k \geq 0} a_k e^{ikx}.
\end{align}
\clearpage
\begin{align}
  \intertext{Let's just consider the diffusion equation ($n=1$). Then the end state is given by}
  u'(x, T) &= \sum_{k\geq 0} a_k \exp(ikx-Tb_1k^2) \label{diffsoln}\\
  \intertext{The nominal objective functional is given by}
  \davg{|u'(x, T)|^2} &= \pi \sum_{k \geq 0} \Big[a_k^2\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, \Big] \label{nomobjdiff}.
  \intertext{Equation (\ref{nomobjdiff}) gives us the elliptical heatmaps we observed previously, as shown in Figure \ref{diff_ellipse}.}
\end{align}

@@

\begin{align}
  \intertext{Let's take a second-order spatial derivative of (\ref{diffsoln})}
  \partial_x^2 u'(x, T) &= \sum_{k\geq 0} -k^2 a_k^2 \exp(ikx-Tb_1k^2)\\
  \intertext{In general, applying a spatial derivative of order $2m$ to (\ref{diffsoln}), yields}
  \partial_x^{2m} u'(x, T) &= \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
  \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical. Consider}
\end{align}
\begin{tiny}
\begin{align*}
  H[u'(T)] &= \davg{|u'(x, T) - (Tb_1)\partial_x^2 u'(x, T) + \frac{(Tb_1)^2}{2!}\partial_x^4 u'(x, T) - \frac{(Tb_1)^3}{3!}\partial_x^6 u'(x, T) ...|^2} \\
  &= \davg{\Big|\sum_{k\geq 0}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}(Tb_1k^2)a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^2}{2!}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^3}{3!}a_k \exp(ikx-Tb_1k^2) + ...\Big|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[a_k \exp(ikx-Tb_1k^2)\Big(1 + (Tb_1k^2) + \frac{(Tb_1k^2)^2}{2!} + \frac{(Tb_1k^2)^3}{3!} ...\Big)\Big]|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx-Tb_1k^2) \Big( \exp(Tb_1k^2) \Big)\Big]|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx)  \Big)\Big]|^2} \\
  &= \pi \sum_{k\geq 0} a_k^2
  % &= \pi \sum_{k \geq 0} \Big[\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, a_k^2\Big].
\end{align*}
\end{tiny}
That gives us spherical objective levelsets. Notice what happens as we increment the number of terms in the above Taylor Series

@@

>>>>

Time-reversible wave equations of the form
\begin{align}
  \partial_t u &= \sum_{n \geq 1} c_n \partial^{2n - 1}_x u \label{waveeq}
  \intertext{are trivial cases of this optimization problem, because the quantity $\davg{u^2(x, t)}$ is conserved. 
  This implies that the objective functional's levelsets form sphere's in $u'$-space, because its value is given by $\davg{u'^2(T)} = \davg{u'^2(0)}$.
  The spatial derivatives in (\ref{waveeq}) are anti-self-adjoint (i.e. $L[u] = -L^{\dagger}[u]$), so the path of steepest descent is always in the radial derivation }
  \frac{\delta H}{\delta u(x, 0)} &= -u'(x, 0) \nonumber
  \intertext{In contrast, diffusive equations of the form}
  \partial_t u &= \sum_{n \geq 1} b_n \partial^{2n}_x u \label{waveeq}
  \intertext{do not conserve $\davg{u^2(x, t)}$, and consequently their objective functionals depend on $u'(x, 0)$'s \textit{direction} along with its magnitude. Suppose initial deviation can be written as}
  u'(x, 0) &= \sum_{k \geq 0} a_k e^{ikx}.
\end{align}
\clearpage
\begin{align}
  \intertext{Let's just consider the diffusion equation ($n=1$) with diffusivity $b_1$. The end state is given by}
  u'(x, T) &= \sum_{k\geq 0} a_k \exp(ikx-Tb_1k^2) \label{diffsoln}\\
  \intertext{and the nominal objective functional is given by}
  \davg{|u'(x, T)|^2} &= \pi \sum_{k \geq 0} \Big[a_k^2\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, \Big] \label{nomobjdiff}.
  \intertext{Equation (\ref{nomobjdiff}) gives us the elliptical heatmaps we observed previously, as shown in Figure \ref{diff_ellipse}.}
\end{align}

@@

\begin{align}
  \intertext{Let's take a second-order spatial derivative of (\ref{diffsoln})}
  \partial_x^2 u'(x, T) &= \sum_{k\geq 0} -k^2 a_k^2 \exp(ikx-Tb_1k^2)\\
  \intertext{In general, applying a spatial derivative of order $2m$ to (\ref{diffsoln}), yields}
  \partial_x^{2m} u'(x, T) &= \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
  \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical.} 
  \intertext{Consider the following objective functional}
\end{align}
\vspace{-1cm}
\begin{tiny}
\begin{align*}
  H[u'(T)] &= \davg{|u'(x, T) - (Tb_1)\partial_x^2 u'(x, T) + \frac{(Tb_1)^2}{2!}\partial_x^4 u'(x, T) - \frac{(Tb_1)^3}{3!}\partial_x^6 u'(x, T) ...|^2} \\
  &= \davg{\Big|\sum_{k\geq 0}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}(Tb_1k^2)a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^2}{2!}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^3}{3!}a_k \exp(ikx-Tb_1k^2) + ...\Big|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[a_k \exp(ikx-Tb_1k^2)\Big(1 + (Tb_1k^2) + \frac{(Tb_1k^2)^2}{2!} + \frac{(Tb_1k^2)^3}{3!} ...\Big)\Big]|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx-Tb_1k^2) \Big( \exp(Tb_1k^2) \Big)\Big]|^2} \\
  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx)  \Big)\Big]|^2} \\
  &= \pi \sum_{k\geq 0} a_k^2
  % &= \pi \sum_{k \geq 0} \Big[\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, a_k^2\Big].
\end{align*}
\end{tiny}
That gives us spherical objective levelsets. Notice what happens as we increment the number of terms in the above Taylor Series

@@

\begin{figure*}[h]
  \centering
  \includegraphics[width=4.5in]{SPHRtest_Nts5a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
\end{figure*}
